---
title: 'Reproducible Research: Peer Assessment 1'
author: 'Vladimir Loiterstein'
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data
Date column is converted to Date type

```{r}
options("scipen"=100, "digits"=4)

df_act<-read.csv("activity.csv",sep=',',header=TRUE)
df_act$date<-as.Date(df_act$date)
```


## What is mean total number of steps taken per day?
###  Total number of steps taken per day
```{r}
df_act_sum<-aggregate(steps ~ date,data=df_act,sum)

df_act_sum
```


### Histogram of the total number of steps taken each day
```{R}
df_hist<-hist(df_act_sum$steps, breaks = 20,main = "Total number of steps taken each day",xlab = "Number of 
steps")
```

### Mean and median of the total number of steps taken per day
```{r,results='hide'}
v_act_mean<-mean(df_act_sum$steps)
v_act_median<-median(df_act_sum$steps)
```

 - Mean is `r v_act_mean`.
 - Meadian is `r v_act_median`.



## What is the average daily activity pattern?
```{r,results='hide'}
df_day_pattern<-aggregate(steps ~ interval,data=df_act,mean)
plot(steps ~ interval,data=df_day_pattern,type="l")
v_max_steps_interval<-df_day_pattern[which.max(df_day_pattern$steps),1]

```
5-minute interval `r v_max_steps_interval` contains the maximum number of steps


## Imputing missing values
Total number of missing values is `r sum(is.na(df_act))`

Replacing NA values with average values for particular 5-min interval.
```{r}
df_empty<-df_act[is.na((df_act$steps)),]
df_filled<-merge(df_empty,df_day_pattern,"interval")
colnames(df_filled)[4]<-"steps"
df_act_new<-rbind(df_act[!is.na((df_act$steps)),],df_filled[,c(4,3,1)])
df_act_new$steps<-round(df_act_new$steps)
df_act_new<-df_act_new[order(df_act_new$date,df_act_new$interval),]
```

df_act_new data set contains missing data filled in


### Histogram of the total number of steps taken each day with filled in NA values.
```{r}
df_act_new_sum<-aggregate(steps ~ date,data=df_act_new,sum)
df_hist<-hist(df_act_new_sum$steps, breaks = 20,main = "Total number of steps taken each day",xlab = "Number of steps")

```


### The mean and median total number of steps taken per day

- Mean is `r mean(df_act_new_sum$steps)`
- Median is `r median(df_act_new_sum$steps)`

As we see values fro Meand and Median decreased afet filling in missed values.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library("lattice")
df_act_new$weekday<-as.factor(
        ifelse(weekdays(df_act_new$date) %in% c("Sunday","Saturday"),"weekend","weekday"))

df_act_weekday_pattern<-aggregate(steps ~ weekday+interval,data=df_act_new,mean)
  
xyplot(steps ~interval|weekday,data=df_act_weekday_pattern,type='l',layout = c(1,2))

```



